---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---


```{r}

library(RMySQL)

mydb = dbConnect(MySQL(), user='root', password='Why940308.', dbname='cs5200', host='127.0.0.1')

```
#data preprocessing

```{sql connection=mydb}
SET GLOBAL local_infile = true;
```


```{r}
data = read.csv(file = 'BirdStrikesData.csv')
head(data,20)
```
#change impact to boolean type
```{r}
data1  = data.frame(data$Record.ID,data$FlightDate,data$Origin.State,data$Aircraft..Airline.Operator,data$Aircraft..Make.Model,data$When..Phase.of.flight,data$Effect..Indicated.Damage,data$Conditions..Sky)


a = nrow(data1['data.Effect..Indicated.Damage'])

count = 0
for(i in 1:a){
  if (data1[i,'data.Effect..Indicated.Damage'] == "No damage"){
    data1[i,'data.Effect..Indicated.Damage'] = 0
  }
  else{
    data1[i,'data.Effect..Indicated.Damage']  = 1
  }
}
head(data1)

```
# change cond as 1,2,3 be like a foreign key
```{r}

b = nrow(data1['data.Conditions..Sky'])
print(b)
for(i in 1:b){
  if(data1[i,'data.Conditions..Sky'] == 'No Cloud'){
    data1[i,'data.Conditions..Sky'] = 1
  }
  else if(data1[i,'data.Conditions..Sky'] == "Some Cloud"){
    data1[i,'data.Conditions..Sky'] =2
    
  }else{
    data1[i,'data.Conditions..Sky'] = 3
  }
}
head(data1)
```


#change column name same with table Incidents.
```{r}

colnames(data1) <- c(  'iid' ,
  'time',
  'origin' ,
  'airline' ,
  'aircraft' ,
  'flightPhase', 
  'impact',
  'cond')


```

#Unified phases
```{r}

data11 = data1

unique(data11["flightPhase"])
e = nrow(data11["flightPhase"])
for(i in 1:e){
  if(data11[i,"flightPhase"]== "Take-off run"){
    data11[i,"flightPhase"] = "takeoff"
  }
  else if(data11[i,"flightPhase"]==	"Landing Roll"){ 
    data11[i,"flightPhase"] = "landing"
  }
  else if (data11[i,"flightPhase"]=="Climb" || data11[i,"flightPhase"]=="Descent" || data11[i,"flightPhase"]=="Approach" ){
    data11[i,"flightPhase"] = "inflight"
  }
  else{
    data11[i,"flightPhase"] = "unknown"
  }
}

head(data11)

```
#set table for conditions.
```{r}
data2  = data.frame(data$Conditions..Sky)
data3 = unique(data2["data.Conditions..Sky"])

data3$cid <- c("1","2","3")
data3$explanation  <- 'none'



```

#change data name as table conditions.
```{r}
colnames(data3) <- c(  'condition' ,
  'cid',
  'explanation')


write.csv(data3,"data3.csv", row.names = FALSE)
```

#creat aid and form the table Airports.
```{r}
data4 = data.frame(data$Airport..Name,data$Origin.State)
head(data4)

data5 = data4[!duplicated(data4["data.Airport..Name"]),]
data5$id = 1:nrow(data5)
data5$code = "none"


```

#out put csv for Airports table. 
```{r}
colnames(data5) <- c( 'airportName' ,
  'state',
  'aid',
  'airportCode')
head(data5)
write.csv(data5,"data5.csv", row.names = FALSE)
```
#set up aid to Incidents table. Take long time to run!
```{r}
datan = data

c = nrow(datan['Airport..Name'])
d = nrow(data5['airportName'])
datan$aid = "none"
for(i in 1:c){
  for(j in 1:d){
    if(datan[i,'Airport..Name']==data5[j,'airportName']){
      datan[i,'aid'] = data5[j,"aid"]
    }
  }
}
head(datan)
```

```{r}
data11$aid = datan[,"aid"]
head(data11)
data12 = data11
```
```{r}


data12 = data11

for(i in 1:nrow(data12['time'])){
  if(data12[i,"time"] != ""){
    f = data12[i,"time"]

    m <- strsplit(f," ",fixed=T)

    y = matrix(unlist(m), ncol=2, byrow=TRUE)

    data12[i,"time"] = y[1,1]
    
  }

}

head(data12)
```




#out put csv for Incidents table.
```{r}
write.csv(data12,"data1.csv", row.names = FALSE)
```


#create tables

```{sql connection=mydb}
drop table if exists Airports

```
```{sql connection=mydb}
drop table if exists conditions

```
```{sql connection=mydb}
drop table if exists Incidents

```

```{sql connection=mydb}
create table Airports(
  aid int not null,
  airportName Text ,
  airportCode Text,
  state text ,
  primary key (aid)
  
)

```

```{sql connection=mydb}
create table conditions(
  `condition` Text,
  cid int not null,
  explanation Text,
  primary key (cid)
  
);

```

```{sql connection=mydb}
create table Incidents(
  iid int not null,
  aid int,
  time date,
  origin Text,
  airline Text,
  aircraft Text,
  flightPhase Text,
  impact BOOLEAN,
  cond int,
  primary key (iid),
  foreign key(aid) references Airports(aid),
  foreign key(cond) references conditions(cid)
);

```




# inster csv to Tables.

```{r}


historydata1<-read.csv("data1.csv",header=TRUE,sep=",")
dbWriteTable(mydb,name = "Incidents",data12,overwrite=TRUE,row.names=FALSE)

historydata2<-read.csv("data3.csv",header=TRUE,sep=",")
dbWriteTable(mydb,"conditions",historydata2,overwrite =TRUE,row.names=FALSE)

historydata3<-read.csv("data5.csv",header=TRUE,sep=",")
dbWriteTable(mydb,"Airports",historydata3,overwrite =TRUE,row.names=FALSE)
```



```{sql connection=mydb}
drop view if exists IncidentsV
```
```{sql connection=mydb}
drop view if exists condV
```
```{sql connection=mydb}
drop view if exists AirportV
```
#3. take look for tables.
```{sql connection=mydb}
CREATE VIEW IncidentsV AS
SELECT *
FROM
Incidents
```

```{sql connection=mydb}
select * from Incidents

```
```{sql connection=mydb}
CREATE VIEW condV AS
SELECT *
FROM
conditions
```

```{sql connection=mydb}
select * from condV
```

```{sql connection=mydb}
CREATE VIEW AirportV AS
SELECT *
FROM
Airports
```
```{sql connection=mydb}
select * from AirportV

```
# remove MILITARY from Incidents table.
```{sql connection=mydb}
DELETE FROM Incidents WHERE airline='MILITARY';

```
```{sql connection=mydb}
select count(airline) from Incidents where airline='MILITARY'

```

4. query for number of strike by different floghtPhase
```{sql connection=mydb}
select flightPhase, count(iid) as num from Incidents group by flightPhase
```

5.above Average 
```{sql connection=mydb}
select flightPhase, count(iid) as aboveAverage from Incidents group by flightPhase having aboveAverage > (select AVG(aboveAverage) from (select flightPhase, count(iid) as aboveAverage from Incidents group by flightPhase having aboveAverage)as f);

```



```{sql connection=mydb}
select * from Incidents
```



# change "time" to date type.

```{sql connection=mydb}
ALTER TABLE Incidents ADD (times DATE);

```


```{sql connection=mydb}
UPDATE Incidents
SET times = str_to_date( time, '%m/%d/%Y' ) where time != '';
```

```{sql connection=mydb}
ALTER TABLE Incidents
DROP COLUMN time;

```

```{sql connection=mydb}
select * from Incidents
```

6.avg for month

```{sql connection=mydb}
select MONTH(times) as months,round(count(iid)/(select count(DISTINCT year(times)) from Incidents )) as Average from Incidents where MONTH(times) != "NA" group by MONTH(times) order by MONTH(times)

```





7.bar plot 
```{r}
df <- dbGetQuery(mydb, "select YEAR(times) as years, count(iid) as counts  from Incidents where YEAR(times) between 2004 and 2012 group by YEAR(times)")

print(df)
```

```{r}
library(ggplot2)
ggplot(data=df, aes(x=years, y=counts)) +
  geom_bar(stat="identity")
```

8.create a table keep who add new data
```{sql connection=mydb}
drop table if exists users
```

```{sql connection=mydb}
DROP TRIGGER IF EXISTS useradd;
```

```{sql connection=mydb}
create table users(
  iid int not null,
  userName text not null,
  primary key(iid)
)


```
aftet insert on Incidents  record who did this change.
```{sql connection=mydb}
create trigger useradd after insert on Incidents 
for each row
BEGIN
insert into users(iid,userName) values (new.iid,(select USER()));
END;
```

```{sql connection=mydb}
insert into Incidents(iid,origin,airline,aircraft,flightPhase,impact,cond,aid,times) VALUES (99999,"New York", "US AIRWAYS*","	B-737-400","takeoff",1,2,160,"2022-3-9")

```

#output.
```{sql connection=mydb}
select * from users

```

```{sql connection=mydb}
select * from Incidents

```




```{r}
dbDisconnect(mydb)
```