---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}
library(RMySQL)


db_user <- 'Jarvis'
db_password <- 'Why940308.'
db_name <- 'P1DB'
db_host <- 'ds5200.cofuluukbiza.us-east-1.rds.amazonaws.com'
db_port <- 3306

mydb <-  dbConnect(RMySQL::MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
```
```{r}
library(RSQLite)

fpath = "/Users/hongyangwang/Library/CloudStorage/OneDrive-Personal/2022northeastern/DS5200/R/Rsqlite5/"
dbfile = "P1DB.sqlite"

mydb <- dbConnect(RSQLite::SQLite(),paste0(fpath,dbfile))
```


#data preprocessing
```{r}
data = read.csv(file = 'BirdStrikesData.csv')
head(data,20)
```
#change impact to boolean type
```{r}
data1  = data.frame(data$Record.ID,data$FlightDate,data$Origin.State,data$Aircraft..Airline.Operator,data$Aircraft..Make.Model,data$When..Phase.of.flight,data$Effect..Indicated.Damage,data$Conditions..Sky)
print(data1)

a = nrow(data1['data.Effect..Indicated.Damage'])
print(a)
count = 0
for(i in 1:a){
  if (data1[i,'data.Effect..Indicated.Damage'] == "No damage"){
    data1[i,'data.Effect..Indicated.Damage'] = 0
  }
  else{
    data1[i,'data.Effect..Indicated.Damage']  = 1
  }
}
print(data1)

```
# change cond as 1,2,3 be like a foreign key
```{r}

b = nrow(data1['data.Conditions..Sky'])
print(b)
for(i in 1:b){
  if(data1[i,'data.Conditions..Sky'] == 'No Cloud'){
    data1[i,'data.Conditions..Sky'] = 1
  }
  else if(data1[i,'data.Conditions..Sky'] == "Some Cloud"){
    data1[i,'data.Conditions..Sky'] =2
    
  }else{
    data1[i,'data.Conditions..Sky'] = 3
  }
}
print(data1)
```


#change column name same with table Incidents.
```{r}

colnames(data1) <- c(  'iid' ,
  'time',
  'origin' ,
  'airline' ,
  'aircraft' ,
  'flightPhase', 
  'impact',
  'cond')


```

#Unified phases
```{r}

data11 = data1

unique(data11["flightPhase"])
e = nrow(data11["flightPhase"])
for(i in 1:e){
  if(data11[i,"flightPhase"]== "Take-off run"){
    data11[i,"flightPhase"] = "takeoff"
  }
  else if(data11[i,"flightPhase"]==	"Landing Roll"){ 
    data11[i,"flightPhase"] = "landing"
  }
  else if (data11[i,"flightPhase"]=="Climb" || data11[i,"flightPhase"]=="Descent" || data11[i,"flightPhase"]=="Approach" ){
    data11[i,"flightPhase"] = "inflight"
  }
  else{
    data11[i,"flightPhase"] = "unknown"
  }
}

print(data11)
print(data1)
```
#set table for conditions.
```{r}
data2  = data.frame(data$Conditions..Sky)
data3 = unique(data2["data.Conditions..Sky"])

data3$cid <- c("1","2","3")
data3$explanation  <- 'none'
print(data3)


```

#change data name as table conditions.
```{r}
colnames(data3) <- c(  'condition' ,
  'cid',
  'explanation')

print(data3)
write.csv(data3,"data3.csv", row.names = FALSE)
```

#creat aid and form the table Airports.
```{r}
data4 = data.frame(data$Airport..Name,data$Origin.State)
print(data4)

data5 = data4[!duplicated(data4["data.Airport..Name"]),]
data5$id = 1:nrow(data5)
data5$code = "none"
print(data5)

```

#out put csv for Airports table. 
```{r}
colnames(data5) <- c( 'airportName' ,
  'state',
  'aid',
  'airportCode')
print(data5)
write.csv(data5,"data5.csv", row.names = FALSE)
```
#set up aid to Incidents table. Take long time to run!
```{r}
datan = data

c = nrow(datan['Airport..Name'])
d = nrow(data5['data.Airport..Name'])
datan$aid = "none"
for(i in 1:c){
  for(j in 1:d){
    if(datan[i,'Airport..Name']==data5[j,'data.Airport..Name']){
      datan[i,'aid'] = data5[j,"id"]
    }
  }
}
print(datan)
```

```{r}
data11$aid = datan[,"aid"]
print(data11)

```
```{r}
print(data11)



for(i in 1:nrow(data11['time'])){
  if(data11[i,"time"] != ""){
    f = data11[i,"time"]

    m <- strsplit(f,"/| ",fixed=T)

    y = matrix(unlist(m), ncol=4, byrow=TRUE)

    data11[i,"time"] = paste(y[1,3],".",y[1,1],".",y[1,2])
    
  }

}
```

```{r}
print(data11)
```


#out put csv for Incidents table.
```{r}
write.csv(data11,"data1.csv", row.names = FALSE)
```


#create tables

```{sql connection=dbcou}
drop table if exists Incidents

```

```{sql connection=mydb}
create table Incidents(
  iid Text not null,
  aid Text not null,
  time date FORMAT 'yyyy-mm-dd',
  origin Text,
  airline Text,
  aircraft Text,
  flightPhase Text,
  impact boolean ,
  cond Text not null,
  primary key (iid),
  foreign key(aid) references Airports(aid)
)

```



```{sql connection=mydb}
create table Airports(
  aid Text not null,
  airportName Text ,
  airportCode Text,
  state text ,
  primary key (aid)
  
)

```

```{sql connection=mydb}
create table conditions(
  cid Text not null,
  condition Text ,
  explanation Text,
  primary key (cid),
  foreign key(cid) references Incidents(cond)
)

```

# inster csv to Tables.

```{r}

historydata<-read.csv("data1.csv",header=TRUE,sep=",")
dbWriteTable(mydb,"Incidents",historydata,append=TRUE,row.names=FALSE)


#historydata<-read.csv("data3.csv",header=TRUE,sep=",")
#dbWriteTable(mydb,"conditions",historydata,append=TRUE,row.names=FALSE)

#historydata<-read.csv("data5.csv",header=TRUE,sep=",")
#dbWriteTable(mydb,"Airports",historydata,append=TRUE,row.names=FALSE)
```



```{sql connection=mydb}
drop view if exists IncidentsV
```

#take look for tables.
```{sql connection=mydb}
CREATE VIEW IncidentsV AS
SELECT *
FROM
Incidents
```

```{sql connection=mydb}
select * from Incidents

```
```{sql connection=mydb}
CREATE VIEW condV AS
SELECT *
FROM
conditions
```

```{sql connection=mydb}
select * from condV
```

```{sql connection=mydb}
CREATE VIEW AirportV AS
SELECT *
FROM
Airports
```
```{sql connection=mydb}
select * from AirportV

```
#remove MILITARY from Incidents table.
```{sql connection=mydb}
DELETE FROM Incidents WHERE airline='MILITARY';

```
```{sql connection=mydb}
select count(airline) from Incidents where airline='MILITARY'

```

4. query for number of strike by different floghtPhase
```{sql connection=mydb}
select flightPhase, count(iid) as num from Incidents group by flightPhase
```

5.above Average 
```{sql connection=mydb}
select flightPhase, count(iid) as aboveAverage from Incidents group by flightPhase having aboveAverage > (select AVG(aboveAverage) from (select flightPhase, count(iid) as aboveAverage from Incidents group by flightPhase having aboveAverage));

```
```{sql connection=mydb}
SELECT time FROM Incidents WHERE time like "" 

```
```{r}
dbDisconnect(mydb)
```